Insurance choice, decision assistance, and insurer steering

### 2021-03-23 finals/lunch-and-learn/lunch-learn-202103.Rmd
update presentation

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble()
xaringanExtra::use_panelset()
```

# Insurance Choice is Confusing

@@

# Insurance Choice is Confusing
.center[
  ![](health-affairs-2017b.png)
]

@@

# Insurance Choice is Confusing
.center[
  ![](health-affairs-2017.png)
]

@@

# Insurance Choice is Confusing

@@

---
# Choice of plans has big consequences

>>>>

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```

# Two motivating facts:

1. Insurance choice is confusing, particularly on the ACA exchanges

2. Size of the market introduces decision problems on a massive scale

---
# 1. Insurance choice is confusing

@@

# 1. Insurance choice is confusing
.center[
  ![](health-affairs-2017b.png)
]

@@

# 1. Insurance choice is confusing

.center[
  ![](health-affairs-2017.png)
]

@@

# 1. Insurance choice is confusing

@@

---
# 2. Potential scale

- 12.7 million enrollees on exchanges in 2016
- Down to 11.4 million in 2020
- Over 27 million remain uninsured

# 2. Potential scale

.pull-left[
Exchanges relatively large, but...

- Large number of people still uninsured
- Many eligible for some type of subsidy on exchanges
]

.pull-right[
  ![:scale 150%](kff-uninsured.png)
]

.footnote[
  [Kaiser Family Foundation, November, 2020](https://www.kff.org/policy-watch/millions-of-uninsured-americans-are-eligible-for-free-aca-health-insurance/)
]

---
# This paper

Lots of people participating in a very complicated decision. With this in mind, we ask the following questions:

1. Does some form of decision assistance "help" in plan selection?
2. Does the type of assistance matter (private versus public)?
3. Do insurers use brokers/agents to steer people to different plans?
4. What are the welfare effects of any such steering?

---
# Outline of talk

1. [Background on the ACA](#background)
2. [Data](#data)
3. [Dominated choices](#dominated)
4. [Effects of decision assistance](#decision-assist)
5. [Insurer steering](#steering)

---
class: inverse, center, middle
name: background

# Background on the ACA exchanges

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
class: clear, center, middle

![:scale 80%](https://media.giphy.com/media/Awb1k8lX8a3Re/giphy.gif)
